maze-dataset

### 2025-03-30 docs/paper/paper.tex
added a bunch of links

  \begin{itemize}
  \tightlist
  \item
    \texttt{RasterizedMazeDataset}, utilizing \texttt{as\_pixels()},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
    similar to the outputs of Németh
    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
  \item
    \texttt{as\_ascii()} provides a format similar to that used in
    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \texttt{MazePlot} provides a feature-rich plotting utility with
    support for multiple paths, heatmaps over positions, and more. This
    is similar to the outputs of
    (\protect\hyperlink{ref-mazegenerator-net}{Alance AB, 2019};
    \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by \texttt{SolvedMaze(...).as\_tokens()} is
  similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \& Wu,
  2023}), but provides over 5.8 million unique formats for converting
  mazes to a text stream, detailed in
  \autoref{tokenized-output-formats}.
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
  this reliably.
\item
  Storing mazes as images is not only difficult to work with, but also
  inefficient. We use an extremely efficient method detailed in
  \autoref{implementation}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
  2008}) appear to simply no longer be accessible.
\end{itemize}

@@

To create a dataset, we first create a \texttt{MazeDatasetConfig}
configuration object, which specifies the seed, number, and size of
mazes, as well as the generation algorithm and its corresponding
parameters. This object is passed to a \texttt{MazeDataset} class to
create a dataset. Crucially, this \texttt{MazeDataset} mimics the
interface of a PyTorch (\protect\hyperlink{ref-pytorch}{Paszke et al.,
2019}) \texttt{Dataset}, and can thus be easily incorporated into
existing data pre-processing and training pipelines, e.g., through the
use of a \texttt{Dataloader} class.

@@

When initializing mazes, further configuration options can be specified
through the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
factory method as necessary. Options allow for saving/loading existing
datasets instead of regenerating, and parallelization options for
generation. Available maze generation algorithms are static methods of
the \texttt{LatticeMazeGenerators} class and include generation
algorithms based on randomized depth-first search, Wilson's algorithm
(\protect\hyperlink{ref-wilson}{Wilson, 1996}), percolation
(\protect\hyperlink{ref-percolation}{Duminil-Copin, 2017};
\protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam, 2004}),
Kruskal's algorithm
(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
others.

@@

All implemented maze generation algorithms are stochastic by nature. For
reproducibility, the \texttt{seed} parameter of
\texttt{MazeDatasetConfig} may be set. In practice, we do not find that
exact duplicates of mazes are generated with any meaningful frequency,
even when generating large datasets.

@@

\begin{figure}[H]
  \centering
  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
    \hline \\[.5em]
    % algorithms
    \texttt{as\_ascii()} & \texttt{as\_pixels()} & \texttt{MazePlot()} \\[.5em]
    % descriptions
      Simple text format for displaying mazes, useful for debugging in a terminal environment.
      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
    \hline \\
    % examples
      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
        \setlength{\baselineskip}{0.9em}
        \input{figures/outputs-ascii-colored.tex} 
      \end{minipage}}
      & \multicolumn{1}{c}{
        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
      }
      & \multicolumn{1}{c}{
        \includegraphics[width=0.25\textwidth, trim={0 0 -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
      } \\[1em]

@@

In previous work, maze tasks have been used with Recurrent Convolutional
Neural Network (RCNN) derived architectures
(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
Huang, et al., 2021}). To facilitate the use of our package in this
context, we replicate the format of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}) and provide the \texttt{RasterizedMazeDataset}
class which returns rasterized pairs of (input, target) mazes as shown
in \autoref{fig:e2h-raster} below.

@@

We convert mazes to token sequences in two steps. First, the maze is
stringified using \texttt{as\_tokens()}. The
\texttt{MazeTokenizerModular} class provides a powerful interface for
configuring maze stringification behavior. Second, the sequence of
strings is tokenized into integers using \texttt{encode()}. Tokenization
uses a fixed vocabulary for simplicity. Mazes up to \(50 \times 50\) are
supported using unique tokens, and up to \(128 \times 128\) when using
coordinate tuple tokens.

There are many algorithms by which one might tokenize a 2D maze into a
1D format usable by autoregressive text models. Training multiple models
on the encodings output from each of these algorithms may produce very
different internal representations, learned solution algorithms, and
levels of performance. To allow exploration of how different maze
tokenization algorithms affect these models, the
\texttt{MazeTokenizerModular} class contains a rich set of options to
customize how mazes are stringified. This class contains 19 discrete
parameters, resulting in over 5.8 million unique tokenizers. But wait,
there's more! There are 6 additional parameters available in the library
which are untested but further expand the the number of tokenizers by a
factor of \(44/3\) to 86 million.

@@

Each \texttt{MazeTokenizerModular} is constructed from a set of several
\texttt{\_TokenizerElement} objects, each of which specifies how
different token regions or other elements of the stringification are
produced.

@@

The tokenizer architecture is purposefully designed such that adding and
testing a wide variety of new tokenization algorithms is fast and
minimizes disturbances to functioning code. This is enabled by the
modular architecture and the automatic inclusion of any new tokenizers
in integration tests. To create a new tokenizer, developers forking the
library may simply create their own \texttt{\_TokenizerElement} subclass
and implement the abstract methods. If the behavior change is
sufficiently small, simply adding a parameter to an existing
\texttt{\_TokenizerElement} subclass and updating its implementation
will suffice. For small additions, simply adding new cases to existing
unit tests will suffice.

The breadth of tokenizers is also easily scaled in the opposite
direction. Due to the exponential scaling of parameter combinations,
adding a small number of new features can significantly slow certain
procedures which rely on constructing all possible tokenizers, such as
integration tests. If any existing subclass contains features which
aren't needed, a developer tool decorator is provided which can be
applied to the unneeded \texttt{\_TokenizerElement} subclasses to prune
those features and compact the available space of tokenizers.

@@

In order to replicate the exact dataset distribution of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}), we allow placing additional constraints in
\texttt{MazeDatasetConfig.endpoint\_kwargs:\ EndpointKwargsType}, such
as enforcing that the start or end point be in a ``dead end'' with only
one accessible neighbor cell. However, combining this with cyclic mazes
(such as those generated with percolation), as was required for the work
in (\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et
al., 2024}), can lead to an absence of valid start and end points.
Placing theoretical bounds on this success rate is difficult, as it
depends on the exact maze generation algorithm and parameters used.
However, our package provides a way to estimate the success rate of a
given configuration using a symbolic regression model trained with PySR
(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
be found in
\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.

Using the estimation simply requires the user to call
\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}
given their predefined \texttt{cfg}.

@@

However, the outputs of this function are not directly usable due to
minor divergences at the endpoints with respect to the percolation
probability \(p\). Since we know that maze success is either guaranteed
or impossible for \(p=0\) and \(p=1\), we define the \texttt{soft\_step}
function to nudge the raw output of the symbolic regression. This
function is defined with:

@@

We combine these to get the \texttt{soft\_step} function, which is
identity-like for \(p \approx 0.5\), and pushes pushes \(x\) to extremes
otherwise. \[
  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
\]

@@

This package utilizes a simple, efficient representation of mazes. Using
an adjacency list to represent mazes would lead to a poor lookup time of
whether any given connection exists, while using an adjacency matrix
would waste memory by failing to exploit the structure (e.g., only 4 of
the diagonals would be filled in). Instead, we describe mazes with the
following simple representation: for a \(d\)-dimensional lattice with
\(r\) rows and \(c\) columns, we initialize a boolean array
\(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
a \texttt{connection\_list}. The value at \(A[0,i,j]\) determines
whether a downward connection exists from node \([i,j]\) to
\([i+1, j]\). Likewise, the value at \(A[1,i,j]\) determines whether a
rightwards connection to \([i, j+1]\) exists. Thus, we avoid duplication
of data about the existence of connections, at the cost of requiring
additional care with indexing when looking for a connection upwards or
to the left. Note that this setup allows for a periodic lattice.

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them. The endpoint
selection can be affected by
\texttt{MazeDatasetConfig.endpoint\_kwargs:\ EndpointKwargsType}, and
complications caused by this are detailed in
\autoref{success-rate-estimation}.

>>>>

  \begin{itemize}
  \tightlist
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}},
    utilizing
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_pixels}{\texttt{as\_pixels()}},
    which can exactly mimic the outputs provided in
    \texttt{easy-to-hard-data}(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild,
    Borgnia, Gupta, Bansal, et al., 2021}) and can be configured to be
    similar to the outputs of Németh
    (\protect\hyperlink{ref-gh_Nemeth_2019}{2019})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    provides a format similar to that used in
    (\protect\hyperlink{ref-gh-oppenheimj2018maze}{Oppenheim, 2018};
    \protect\hyperlink{ref-eval-gpt-visual}{Singla, 2023})
  \item
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html\#MazePlot}{\texttt{MazePlot}}
    provides a feature-rich plotting utility with support for multiple
    paths, heatmaps over positions, and more. This is similar to the
    outputs of (\protect\hyperlink{ref-mazegenerator-net}{Alance AB,
    2019}; \protect\hyperlink{ref-gh_Ehsan_2022}{Ehsan, 2022};
    \protect\hyperlink{ref-mathematica-maze}{Guo et al., 2011};
    \protect\hyperlink{ref-mdl-suite}{Nag, 2020})
  \end{itemize}
\item
  The text format provided by
  \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{SolvedMaze(...).as\_tokens()}}
  is similar to that of (\protect\hyperlink{ref-eval-LLM-graphs}{Liu \&
  Wu, 2023}), but provides over 5.8 million unique formats for
  converting mazes to a text stream, detailed in
  \autoref{tokenized-output-formats}.
\item
  For rigorous investigations of the response of a model to various
  distributional shifts, preserving metadata about the generation
  algorithm with the dataset itself is essential. To this end, our
  package efficiently stores the dataset along with its metadata in a
  single human-readable file (\protect\hyperlink{ref-zanj}{M.
  Ivanitskiy, n.d.}). As far as we are aware, no existing packages do
  this reliably.
\item
  Storing mazes as images is not only difficult to work with, but also
  inefficient. We use an extremely efficient method detailed in
  \autoref{implementation}.
\item
  Our package is easily installable with source code freely available.
  It is extensively tested, type hinted, benchmarked, and documented.
  Many other maze generation packages lack this level of rigor and
  scope, and some (\protect\hyperlink{ref-ayaz2008maze}{Ayaz et al.,
  2008}) appear to simply no longer be accessible.
\end{itemize}

@@

To create a dataset, we first create a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
configuration object, which specifies the seed, number, and size of
mazes, as well as the generation algorithm and its corresponding
parameters. This object is passed to a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
class to create a dataset. Crucially, this
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset}{\texttt{MazeDataset}}
mimics the interface of a PyTorch
(\protect\hyperlink{ref-pytorch}{Paszke et al., 2019})
\href{https://pytorch.org/docs/stable/data.html}{\texttt{Dataset}}, and
can thus be easily incorporated into existing data pre-processing and
training pipelines, e.g., through the use of a \texttt{DataLoader}
class.

@@

When initializing mazes, further configuration options can be specified
through the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.from_config}{\texttt{from\_config()}}
factory method as necessary. Options allow for saving/loading existing
datasets instead of regenerating, and parallelization options for
generation. Available maze generation algorithms are static methods of
the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMazeGenerators}{\texttt{LatticeMazeGenerators}}
class and include generation algorithms based on randomized depth-first
search, Wilson's algorithm (\protect\hyperlink{ref-wilson}{Wilson,
1996}), percolation (\protect\hyperlink{ref-percolation}{Duminil-Copin,
2017}; \protect\hyperlink{ref-percolation-clustersize}{Fisher \& Essam,
2004}), Kruskal's algorithm
(\protect\hyperlink{ref-kruskal1956shortest}{Kruskal, 1956}), and
others.

@@

All implemented maze generation algorithms are stochastic by nature. For
reproducibility, the \texttt{seed} parameter of
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig}{\texttt{MazeDatasetConfig}}
may be set. In practice, we do not find that exact duplicates of mazes
are generated with any meaningful frequency, even when generating large
datasets.

@@

\begin{figure}[H]
  \centering
  \begin{tabular}{p{1.5in} p{1.5in} p{1.5in}} 
    \hline \\[.5em]
    % algorithms
    \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_ascii}{\texttt{as\_ascii()}}
    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset.html#LatticeMaze.as_pixels}{\texttt{as\_pixels()}}
    & \href{https://understanding-search.github.io/maze-dataset/maze_dataset/plotting.html#MazePlot}{\texttt{MazePlot()}} \\[.5em]
    % descriptions
      Simple text format for displaying mazes, useful for debugging in a terminal environment.
      & \texttt{numpy} array of \texttt{dtype=uint8} and shape \texttt{(height, width, 3)}. The last dimension is RGB color.
      & feature-rich plotting utility with support for multiple paths, heatmaps over positions, and more. \\[1em]
    \hline \\
    % examples
      \multicolumn{1}{c}{\begin{minipage}[b]{1.6in}
        \setlength{\baselineskip}{0.9em}
        \input{figures/outputs-ascii-colored.tex} 
      \end{minipage}}
      & \multicolumn{1}{c}{
        \includegraphics[width=0.25\textwidth]{figures/outputs-pixels.pdf}
      }
      & \multicolumn{1}{c}{
        \includegraphics[width=0.25\textwidth, trim={0 0 -.3cm, -.5cm}, clip]{figures/outputs-mazeplot.pdf}
      } \\[1em]

@@

In previous work, maze tasks have been used with Recurrent Convolutional
Neural Network (RCNN) derived architectures
(\protect\hyperlink{ref-deepthinking}{Schwarzschild, Borgnia, Gupta,
Huang, et al., 2021}). To facilitate the use of our package in this
context, we replicate the format of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}) and provide the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/rasterized.html\#RasterizedMazeDataset}{\texttt{RasterizedMazeDataset}}
class which returns rasterized pairs of (input, target) mazes as shown
in \autoref{fig:e2h-raster} below.

@@

We convert mazes to token sequences in two steps. First, the maze is
stringified using
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDataset.as_tokens}{\texttt{as\_tokens()}}.
The
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
class provides a powerful interface for configuring maze stringification
behavior. Second, the sequence of strings is tokenized into integers
using \texttt{encode()}. Tokenization uses a fixed vocabulary for
simplicity. Mazes up to \(50 \times 50\) are supported using unique
tokens, and up to \(128 \times 128\) when using coordinate tuple tokens.

There are many algorithms by which one might tokenize a 2D maze into a
1D format usable by autoregressive text models. Training multiple models
on the encodings output from each of these algorithms may produce very
different internal representations, learned solution algorithms, and
levels of performance. To allow exploration of how different maze
tokenization algorithms affect these models, the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
class contains a rich set of options to customize how mazes are
stringified. This class contains 19 discrete parameters, resulting in
over 5.8 million unique tokenizers. But wait, there's more! There are 6
additional parameters available in the library which are untested but
further expand the the number of tokenizers by a factor of \(44/3\) to
86 million.

@@

Each
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#MazeTokenizerModular}{\texttt{MazeTokenizerModular}}
is constructed from a set of several
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
objects, each of which specifies how different token regions or other
elements of the stringification are produced.

@@

The tokenizer architecture is purposefully designed such that adding and
testing a wide variety of new tokenization algorithms is fast and
minimizes disturbances to functioning code. This is enabled by the
modular architecture and the automatic inclusion of any new tokenizers
in integration tests. To create a new tokenizer, developers forking the
library may simply create their own
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclass and implement the abstract methods. If the behavior change is
sufficiently small, simply adding a parameter to an existing
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclass and updating its implementation will suffice. For small
additions, simply adding new cases to existing unit tests will suffice.

The breadth of tokenizers is also easily scaled in the opposite
direction. Due to the exponential scaling of parameter combinations,
adding a small number of new features can significantly slow certain
procedures which rely on constructing all possible tokenizers, such as
integration tests. If any existing subclass contains features which
aren't needed, a developer tool decorator
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization/modular/element_base.html\#mark_as_unsupported}{\texttt{@mark\_as\_unsupported}}
is provided which can be applied to the unneeded
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/tokenization.html\#_TokenizerElement}{\texttt{\_TokenizerElement}}
subclasses to prune those features and compact the available space of
tokenizers.

@@

In order to replicate the exact dataset distribution of
(\protect\hyperlink{ref-easy_to_hard}{Schwarzschild, Borgnia, Gupta,
Bansal, et al., 2021}), we allow placing additional constraints in
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
such as enforcing that the start or end point be in a ``dead end'' with
only one accessible neighbor cell. However, combining this with cyclic
mazes (such as those generated with percolation), as was required for
the work in
(\protect\hyperlink{ref-knutson2024logicalextrapolation}{Knutson et al.,
2024}), can lead to an absence of valid start and end points. Placing
theoretical bounds on this success rate is difficult, as it depends on
the exact maze generation algorithm and parameters used. However, our
package provides a way to estimate the success rate of a given
configuration using a symbolic regression model trained with PySR
(\protect\hyperlink{ref-pysr}{Cranmer, 2023}). More details on this can
be found in
\href{https://understanding-search.github.io/maze-dataset/notebooks/estimate_dataset_fractions.html}{\texttt{estimate\_dataset\_fractions.ipynb}}.

Using the estimation simply requires the user to call
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#MazeDatasetConfig.success_fraction_compensate}{\texttt{cfg\_new:\ MazeDatasetConfig\ =\ cfg.success\_fraction\_compensate()}},
providing their initial \texttt{cfg} and then using the returned
\texttt{cfg\_new} in its place.

@@

However, the outputs of this function are not directly usable due to
minor divergences at the endpoints with respect to the percolation
probability \(p\). Since we know that maze success is either guaranteed
or impossible for \(p=0\) and \(p=1\), we define the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
function to nudge the raw output of the symbolic regression. This
function is defined with:

@@

We combine these to get the
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/success_predict_math.html\#soft_step}{\texttt{soft\_step}}
function, which is identity-like for \(p \approx 0.5\), and pushes
pushes \(x\) to extremes otherwise. \[
  \text{soft\_step}(x, p, \alpha, w) = h(x, A(p, \alpha, w))
\]

@@

This package utilizes a simple, efficient representation of mazes. Using
an adjacency list to represent mazes would lead to a poor lookup time of
whether any given connection exists, while using an adjacency matrix
would waste memory by failing to exploit the structure (e.g., only 4 of
the diagonals would be filled in). Instead, we describe mazes with the
following simple representation: for a \(d\)-dimensional lattice with
\(r\) rows and \(c\) columns, we initialize a boolean array
\(A = \{0, 1\}^{d \times r \times c}\), which we refer to in the code as
a
\href{https://understanding-search.github.io/maze-dataset/maze_dataset.html\#LatticeMaze.connection_list}{\texttt{connection\_list}}.
The value at \(A[0,i,j]\) determines whether a downward connection
exists from node \([i,j]\) to \([i+1, j]\). Likewise, the value at
\(A[1,i,j]\) determines whether a rightwards connection to \([i, j+1]\)
exists. Thus, we avoid duplication of data about the existence of
connections, at the cost of requiring additional care with indexing when
looking for a connection upwards or to the left. Note that this setup
allows for a periodic lattice.

To produce solutions to mazes, two points are selected uniformly at
random without replacement from the connected component of the maze, and
the \(A^*\) algorithm (\protect\hyperlink{ref-A_star}{Hart et al.,
1968}) is applied to find the shortest path between them. The endpoint
selection can be affected by
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#MazeDatasetConfig.endpoint_kwargs}{\texttt{MazeDatasetConfig.endpoint\_kwargs:}}
\href{https://understanding-search.github.io/maze-dataset/maze_dataset/dataset/maze_dataset_config.html\#EndpointKwargsType}{\texttt{EndpointKwargsType}},
and complications caused by this are detailed in
\autoref{success-rate-estimation}.
